### 2022-04-27 paper/paper.tex
set up paper things

\documentclass[twocolumn, twocolappendix]{aastex631}
\received{\today}
\shorttitle{Detecting DCOs with LISA}
\graphicspath{{../figures/}}

\usepackage{lipsum}
\usepackage{physics}
\usepackage{multirow}
\usepackage{xspace}
\usepackage{natbib}
\usepackage{fontawesome5}
\usepackage{xcolor}
\usepackage{wrapfig}
\usepackage[figuresright]{rotating}

% remove indents in footnotes
\usepackage[hang,flushmargin]{footmisc}

\newcommand{\todo}[1]{{\color{red}{[TODO: #1}]}}

% custom function for adding units
\makeatletter
\newcommand{\unit}[1]{%
    \,\mathrm{#1}\checknextarg}
\newcommand{\checknextarg}{\@ifnextchar\bgroup{\gobblenextarg}{}}
\newcommand{\gobblenextarg}[1]{\,\mathrm{#1}\@ifnextchar\bgroup{\gobblenextarg}{}}
\makeatother

\begin{document}

\title{The NEOCP in the Era of LSST}

% affiliations
\newcommand{\cfa}{Center for Astrophysics | Harvard \& Smithsonian, 60 Garden Street, Cambridge, MA 02138, USA}
\newcommand{\mpa}{Max-Planck-Institut für Astrophysik, Karl-Schwarzschild-Straße 1, 85741 Garching, Germany}
\newcommand{\cca}{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY, 10010, USA}
\newcommand{\UvA}{Anton Pannekoek Institute for Astronomy \& Grappa, University of Amsterdam, Postbus 94249, 1090 GE Amsterdam, The Netherlands}
\newcommand{\UW}{Department of Astronomy, University of Washington, Seattle, WA, 98195}

\author[0000-0001-6147-5761]{T. Wagg}
\affiliation{\UW}

\correspondingauthor{Tom Wagg}
\email{tomjwagg@gmail.com}

\begin{abstract}
    \input{sections/abstract.tex}
\end{abstract}

\keywords{gravitational waves, gravitational wave detectors, compact objects, stellar mass black holes, neutron stars, binary stars, stellar evolution, pulsars}

\section{Introduction} \label{sec:intro}
\input{sections/introduction.tex}

\section{Method} \label{sec:method}
\input{sections/methods.tex}

\section{Results} \label{sec:results}
\input{sections/results.tex}

\section{Discussion} \label{sec:discussion}
\input{sections/discussion.tex}

\section{Conclusion \& Summary} \label{sec:conclusion}
\input{sections/conclusion.tex}

\begin{acknowledgements}
    We'll thank some people here
\end{acknowledgements}

\software{}

\bibliographystyle{aasjournal}
\bibliography{paper}{}

% \restartappendixnumbering

% \allowdisplaybreaks
% \appendix
% \input{sections/appendix.tex}

\end{document}