### 2022-05-15 paper/paper.tex
start writing the methods

\documentclass[twocolumn, twocolappendix]{aastex631}
\received{\today}
\shorttitle{NEOCP in LSST era}
\graphicspath{{figures/}}

\usepackage{lipsum}
\usepackage{physics}
\usepackage{multirow}
\usepackage{xspace}
\usepackage{natbib}
\usepackage{fontawesome5}
\usepackage{xcolor}
\usepackage{wrapfig}
\usepackage[figuresright]{rotating}

% remove indents in footnotes
\usepackage[hang,flushmargin]{footmisc}

\newcommand{\todo}[1]{{\color{red}{[TODO: #1}]}}
\newcommand{\dig}{\texttt{digest2}}

% custom function for adding units
\makeatletter
\newcommand{\unit}[1]{%
    \,\mathrm{#1}\checknextarg}
\newcommand{\checknextarg}{\@ifnextchar\bgroup{\gobblenextarg}{}}
\newcommand{\gobblenextarg}[1]{\,\mathrm{#1}\@ifnextchar\bgroup{\gobblenextarg}{}}
\makeatother

\begin{document}

\title{The NEOCP in the Era of LSST}

% affiliations
\newcommand{\UW}{Department of Astronomy, University of Washington, Seattle, WA, 98195}

\author[0000-0001-6147-5761]{T. Wagg}
\affiliation{\UW}

\author[0000-0003-1996-9252]{M. Juric}
\affiliation{\UW}

\author{\dots more}

\correspondingauthor{Tom Wagg}
\email{tomjwagg@gmail.com}

\begin{abstract}
    \input{sections/abstract.tex}
\end{abstract}

\keywords{Near-Earth objects, Asteroids, Solar system, Small Solar System bodies, Surveys}

\section{Introduction} \label{sec:intro}
\input{sections/introduction.tex}

\section{Method} \label{sec:method}
\input{sections/methods.tex}

\section{Results} \label{sec:results}
\input{sections/results.tex}

\section{Discussion} \label{sec:discussion}
\input{sections/discussion.tex}

\section{Conclusion \& Summary} \label{sec:conclusion}
\input{sections/conclusion.tex}

\begin{acknowledgements}
    Acknowledgements
\end{acknowledgements}

\software{\texttt{astroML} \citep{VanderPlas+2012,VanderPlas+2014}, \texttt{scipy} \citep{Virtanen+2020}}

\bibliographystyle{aasjournal}
\bibliography{paper}{}

% \restartappendixnumbering

% \allowdisplaybreaks
% \appendix
% \input{sections/appendix.tex}

\end{document}